\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-06-07 nampi.tex
learning from scratch as well as an introduction rewrite

\usepackage{xcolor}
\definecolor{pop1}{HTML}{1f78b4}
\definecolor{pop2}{HTML}{297F23}
\definecolor{pop3}{HTML}{d95f02}
\definecolor{orange}{HTML}{d95f02}
\definecolor{teal}{HTML}{1b9e77}
\newcommand{\pop}[1]{\textcolor{pop1}{#1}}
\newcommand{\popp}[1]{\textcolor{pop2}{#1}}
\newcommand{\tree}[1]{\textcolor{pop3}{#1}}
\newcommand{\orange}[1]{\textcolor{orange}{#1}}
\newcommand{\teal}[1]{\textcolor{teal}{#1}}

@@

\begin{abstract}
  Successful approaches to program induction require a hand-engineered
  domain-specific language (DSL), constraining the space of allowed
  programs and imparting prior knowledge of the domain.  We contribute
  a program induction algorithm called \system that learns a DSL while
  jointly training a neural network to efficiently search for programs
  in the learned DSL.\@ We use our model to synthesize functions on lists,
  edit text, and solve symbolic regression problems, showing how the
  model learns a domain-specific library of program components for
  expressing solutions to problems in the domain.
\end{abstract}

\section{Introduction}

@@

The FlashFill system~\cite{gulwani2011automating} developed by
Microsoft researchers and now embedded in Excel solves problems such
as these and is probably the best known practical program-induction
algorithm, but researchers in programming languages and AI have built
successful program induction algorithms for many applications, 
such as handwriting recognition and generation~\cite{lake2015human},
procedural graphics~\cite{ellis2017learning}, question
answering~\cite{johnson2017clevr} and robot motion
planning~\cite{devlin2017neural}, to name just a few.
These systems work in different ways, but most hinge upon having a
carefully engineered \textbf{Domain Specific Language (DSL)}.  This is
especially true for systems such as FlashFill that aim to induce a
wide range of programs very quickly, in a few seconds or less.  DSLs
constrain the search over programs with strong prior knowledge in the
form of a restricted set of programming primitives tuned to the needs
of the domain: for text editing, these are operations like appending
strings and splitting on characters.

@@

  \end{tabular}
  \\\bottomrule\\
\end{tabular}}
\caption{Top: Tasks from each domain, each followed by the programs \system discovers for them. Bottom: Several examples from learned DSL. Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ in List Functions)}\label{initialExampleDSL}
\end{table*}
Because any computable learning problem can in principle be cast as
program induction, it is important to delimit our focus.  In contrast
to computer assisted programming~\cite{solar2008program} or genetic
programming~\cite{DBLP:books/daglib/0070933}, our goal is not to automate software
engineering, to learn to synthesize large bodies of code, or to learn
complex programs starting from scratch.  Ours is a basic AI goal:
capturing the human ability to learn to think flexibly and efficiently
in new domains --- to learn what you need to know about a domain so you
don't have to solve new problems starting from scratch.  We are
focused on the kinds of problems that humans can solve relatively
quickly, once they acquire the relevant domain expertise.  These
correspond to tasks solved by short programs --- if you have an
expressive DSL.  Even with a good DSL, program search may be
intractable;
%
% , and adding new library routines to the DSL only broadens the
% search space;
%
so we amortize the cost of search by training a
neural network to assist the search procedure.

Our algorithm takes inspiration from several ways that skilled human
programmers have learned to code: Skilled coders build libraries of
reusable subroutines that are shared across related programming tasks,
and can be composed to generate increasingly complex and powerful
subroutines.  In text editing, a good library should support routines
for splitting on characters, but also specialize these routines to
split on particular characters such as spaces or commas that are
frequently used to delimit substrings across tasks.  Skilled coders
also learn to recognize what kinds of programming idioms and library
routines would be useful for solving the task at hand, even if they
cannot instantly work out the details.  In text editing, one might
learn that if outputs are consistently shorter than inputs, removing
characters is likely to be part of the solution; if every output
contains a constant substring (e.g., ``Dr.''), inserting or appending
that constant string is likely to be a subroutine.

Our \system
algorithm incorporates these insights by iterating through three steps. 
The \textbf{Search} step takes a given set of \textbf{tasks}, typically
several hundred, and searches for compact programs that solve these tasks
guided by the current DSL and neural network.
The \textbf{Compress} step grows the library (or DSL) of
domain-specific subroutines which allow the agent to more compactly
write programs in the domain; it modifies the structure of the DSL by
discovering regularities across programs found during search, compressing
them to distill out common code fragments across successful programs.
The \textbf{Compile} step improves the search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}.%%  (CITE Goodman and Stuhlmuller,
%% and others too on amortized inference.)

The learned DSL effectively encodes a prior on programs likely to
solve tasks in the domain, while the neural net looks at the example
input-output pairs for a specific task and produces a ``posterior''
for programs likely to solve that specific task.  The neural network
thus functions as a \textbf{recognition model} supporting a form of
approximate Bayesian program induction, jointly trained with a
\textbf{generative model} for programs encoded in the DSL, in the
spirit of the Helmholtz machine~\cite{hinton1995wake}). The
recognition model ensures that searching for programs remains
tractable even as the DSL (and hence the search space for programs)
expands.

We apply \system to three domains:
list processing; text editing (in the style of FlashFill~\cite{gulwani2011automating}); and symbolic regression.
 For each of these we initially provide a generic
 set of programming primitives.
Our algorithm then constructs
its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).

\begin{comment}
Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}),
neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
which also learns a library of subroutines.
We find EC motivating, and go beyond it and other prior work through the following contributions:%% \begin{itemize}
%%   \setlength\itemsep{0}
%%   \item We show how to learn-to-learn programs in an expressive Lisp-like programming language;
%%   \item We give an algorithm for learning DSLs, built on a formalism known
%%     as Fragment Grammars~\cite{tim}; and 
%%   \item We give a hierarchical Bayesian framing of the problem that allows a
%%     joint inference of the DSL and recognition model.
%% \end{itemize}
\noindent\\\textbf{Contributions.} (1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
(2) We give an algorithm for learning DSLs,
built on a formalism known as Fragment Grammars~\cite{tim};
and (3) We give a hierarchical Bayesian framing of
the problem that
allows joint inference of the DSL and neural recognition model.
  \end{comment}

@@

\section{The \system Algorithm}
\begin{figure}\centering
  \begin{tikzpicture}
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs};
    \node[align = center](DSL) at (3,1) {DSL};
    \node[align = center](recognitionModel) at (9,1) {Neural\\recognition \\model};

    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
  \end{scope}
    \end{tikzpicture}
  \caption{\system solves for programs, the DSL, and a recognition model. Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
\end{figure}

@@

\begin{wrapfigure}{r}{0.55\textwidth}\centering
  \begin{tabular}{lll}
    \toprule
    Name & Input & Output \\\midrule
    repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
    rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
    keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
    product & [7\, 1\, 6\, 2] & 84 \\
    \bottomrule
  \end{tabular}
  \captionof{table}{Some tasks in our list function domain. See the
  supplement for the complete data set.}\label{listexamples}
\end{wrapfigure}
We created 236 human-interpretable list manipulation tasks, each with 15
input/output examples (Tbl.~\ref{listexamples}).
Our data set is interesting in three major ways: many of the tasks require
complex solutions; the tasks were not generated from some latent DSL, and
the agent must learn to solve these complicated problems from only 236
tasks.
Our data set assumes arithmetic operations as well as sequence operations,
so we additionally provide our system with the following arithmetic
primitives: \code{mod}, \code{*}, \code{>}, \code{is-square},
\code{is-prime}.

We evaluated \system on  random 50/50 test/train split.
Interestingly, we found that the recognition model provided little benefit for the training
tasks. However, it yielded faster search times on held out tasks, allowing
more tasks to be solved before timing out.
The system composed 38 new subroutines, yielding a more expressive DSL more closely matching the
domain (left of Tbl.~\ref{initialExampleDSL}, right of
Fig.~\ref{fragmentExample}). See the supplement for a complete list of
DSL primitives discovered by \systemEnding.

@@

Because our enumerative search procedure cannot generate string
constants, we instead enumerate programs with string-valued
parameters.  For example, to learn a program that prepends ``Dr.'', we
enumerate $\text{\code{(}}f_3\code{ string s)}$ -- where $f_3$ is the
learned appending primitive (Fig.~\ref{initialExampleDSL}) --- and then
define $\probability[x|p]$ by approximately marginalizing out the
string parameters via a simple dynamic program.
In Sec.~\ref{regressionSection}, we will use a similar trick to
synthesize programs containing real numbers, but using gradient
descent instead of dynamic programming.

@@

\section{Symbolic Regression: Programs from visual input}\label{regressionSection}
We apply \system
to symbolic regression problems.  Here, the
agent observes points along the curve of a function, and must write a
program that fits those points.  We initially equip our learner with
addition, multiplication, and division, and task it with solving
100 symbolic regression problems, each either a polynomial of degree
1--4 or a rational function.  The recognition model is a
convolutional network that observes an image of the target function's
graph (Fig.~\ref{functions}) --- visually, different kinds of
polynomials and rational functions produce different kinds of graphs,
and so the recognition model can learn to look at a graph and predict
what kind of function best explains it.  A key difficulty, however, is
that these problems are best solved with programs containing real
numbers.  Our solution to this difficulty is to enumerate
 programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent to fit the parameters.
We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.

\begin{wrapfigure}{r}{0.45\textwidth}\vspace{-0.5cm} \newcommand{\functionSize}{1.5cm}\centering
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/48.png}
  \includegraphics[width = \functionSize]{figures/functions/102.png}
  \includegraphics[width = \functionSize]{figures/functions/116.png}\\
  \vspace{1pt}\includegraphics[width = \functionSize]{figures/functions/181.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/149.png}
    \includegraphics[width = \functionSize]{figures/functions/187.png}
  \caption{Recognition model input for symbolic regression. DSL learns subroutines for polynomials (top row) and rational functions (bottom row) while the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines best explains the observation.}\label{functions}\vspace{-2cm}
\end{wrapfigure}
\system learns a DSL containing 13 new functions,
most of which are templates for polynomials of different orders or ratios of polynomials.
It also learns to find programs that minimize the number of continuous degrees of freedom.
For example, it learns to represent linear functions with the program
\code{(* real (+ x real))}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing --- both the implicit bias towards shorter programs and the likelihood model's BIC penalty.

@@

We compare with ablations of our model on held out tasks.
The purpose of this ablation study is 
both to examine the role of each component of \systemEnding,
as well as to compare with
prior approaches in the literature:
a head-to-head
comparison of program synthesizers is complicated by the fact that
each system, including ours, makes idiosyncratic 
assumptions about the space of programs and the statement of tasks.
\begin{wraptable}{r}{0.55\textwidth}
\vspace{-5pt}
%% This radically shrinks the table
\tabcolsep=2.5pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{lcccccc}
  \toprule& Ours& 
    No NN & SE&NPS & PCFG & Enum \\
  \midrule
  \multicolumn{7}{c}{\emph{List Processing}}\\
  \midrule
  \% solved&\textbf{79\%}\footnotemark & 
  76\% &71\%&35\%&62\%&37\%\\
  Solve time&  4.1s&5.8s&10.6s&34.7s&43.4s&20.2s\\
  \midrule
  \multicolumn{7}{c}{\emph{Text Editing}}\\
  \midrule
  \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
  Solve time& 29s&65s&38s&80s&--&235s\\
  \midrule
  \multicolumn{7}{c}{\emph{Symbolic Regression}}\\
  \midrule
  \% solved&   \textbf{84\% }&75\%&62\%&38\%&38\%&37\% \\
  Solve time&  24s& 40s  &28s&31s&55s&29s\\
  \bottomrule
  \end{tabular}
\caption{\% held-out test tasks solved. Solve time: averaged over solved
  tasks.  %SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
  %structure learning.  Enum: model w/o any learning.  %% MDL:
  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
  %% symbolic regression MDL is 0 nats.
  %% For symbolic regression MDL is a
  %% proxy for \# of continuous parameters.
}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
Nevertheless, much prior work can be modeled within our setup. 
We compare with the following ablations (Tbl~\ref{baselineComparisons};
Fig~\ref{learningCurves}):
\\\noindent \textbf{No NN:} lesions the recognition model.
\\\noindent \textbf{NPS}, which does not learn the DSL,
instead learning the recognition model
from samples drawn from the fixed DSL.
We call this NPS (Neural Program Synthesis)
because this is closest to how
RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
\\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL. This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
\\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first search step.
\footnotetext[2]{Due to a last-minute bug we ran the list processing experiments
without training the recognition model on ``Helmholtz machine'' style data. We
are rerunning this experiment and anticipate improved quantitative results for
list processing --- notice the recognition model helps much more for text editing and symbolic regression.} &
%For each domain,
%% We are interested both in how many tasks the
%% agent can solve and how quickly it can find those solutions.
%% Tbl.~\ref{baselineComparisons}
%% compares our model against these alternatives.
%% We consistently
%% improve on the baselines,
%% and find that lesioning the recognition model
%% and lesioning it also slows down the convergence of the algorithm,
%% taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
%% This supports a view of the recognition model as a way of amortizing the cost of search.

>>>>

\usepackage{xcolor}
\definecolor{pop1}{HTML}{1F78b4}
\definecolor{pop2}{HTML}{297F23}
\definecolor{pop3}{HTML}{d95F02}
\definecolor{orange}{HTML}{d95F02}
\definecolor{teal}{HTML}{1b9e77}
\newcommand{\pop}[1]{\textcolor{pop1}{#1}}
\newcommand{\popp}[1]{\textcolor{pop2}{#1}}
\newcommand{\tree}[1]{\textcolor{pop3}{#1}}
\newcommand{\orange}[1]{\textcolor{orange}{#1}}
\newcommand{\teal}[1]{\textcolor{teal}{#1}}

@@

%% \begin{abstract}
%%   Successful approaches to program induction require a hand-engineered
%%   domain-specific language (DSL), constraining the space of allowed
%%   programs and imparting prior knowledge of the domain.  We contribute
%%   a program induction algorithm called \system that learns a DSL while
%%   jointly training a neural network to efficiently search for programs
%%   in the learned DSL.\@ We use our model to synthesize functions on lists,
%%   edit text, and solve symbolic regression problems, showing how the
%%   model learns a domain-specific library of program components for
%%   expressing solutions to problems in the domain.
%% \end{abstract}

\section{Introduction: The Centrality of Domain-Specific Languages}

@@

The FlashFill system~\cite{gulwani2011automating} developed by
Microsoft researchers and now embedded in Excel solves problems such
as these and is probably the best known practical program-induction
algorithm, but researchers in programming languages and AI have built
successful program induction algorithms for many applications, 
such as handwriting recognition and generation~\cite{lake2015human},
procedural graphics~\cite{ellis2017learning}, question
answering~\cite{johnson2017clevr} and robot motion
planning~\cite{devlin2017neural}, to name just a few.
These systems work in different ways, but most hinge upon  a
carefully engineered \textbf{Domain Specific Language (DSL)}.  This is
especially true for systems such as FlashFill that aim to induce a
wide range of programs very quickly, in a few seconds or less.  DSLs
constrain the search over programs with strong prior knowledge in the
form of a restricted set of programming primitives tuned to the needs
of the domain: for text editing, these are operations like appending
strings and splitting on characters.

@@

  \end{tabular}
  \\\bottomrule\\
\end{tabular}}
\caption{Top: Tasks from three domains we apply our algorithm to, each followed by the programs \system discovers for them. Bottom: Several examples from learned DSL. Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}
\end{table*}
Because any computable learning problem can in principle be cast as
program induction, it is important to delimit our focus.  In contrast
to computer assisted programming~\cite{solar2008program} or genetic
programming~\cite{DBLP:books/daglib/0070933}, our goal is not to automate software
engineering, to learn to synthesize large bodies of code, or to learn
complex programs starting from scratch.  Ours is a basic AI goal:
capturing the human ability to learn to think flexibly and efficiently
in new domains --- to learn what you need to know about a domain so you
don't have to solve new problems starting from scratch.  We are
focused on the kinds of problems that humans can solve relatively
quickly, once they acquire the relevant domain expertise.  These
correspond to tasks solved by short programs --- if you have an
expressive DSL.  Even with a good DSL, program search may be
intractable;
%
% , and adding new library routines to the DSL only broadens the
% search space;
%
so we train a
neural network to guide the search procedure.

\section{The \system Algorithm}

We take inspiration from several ways that skilled human
programmers have learned to code: Skilled coders build libraries of
reusable subroutines that are shared across related programming tasks,
and can be composed to generate increasingly complex and powerful
subroutines.  In text editing, a good library should support routines
for splitting on characters, but also specialize these routines to
split on characters such as spaces or commas that are
frequently used to delimit substrings.  Skilled coders
also learn to recognize what kinds of programming idioms and library
routines would be useful for solving the task at hand, even if they
cannot instantly work out the details.  In text editing, one might
learn that if outputs are consistently shorter than inputs, removing
characters is likely to be part of the solution.%% ; if every output
%% contains a constant substring (e.g., ``Dr.''), inserting or appending
%% that constant string is likely to be a subroutine.

Our algorithm is called \system because it is based on a
novel kind of ``wake-sleep'' learning (c.f. \cite{hinton1995wake}), iterating
between ``wake'' and ``sleep'' phases to achieve three goals: finding
programs that solve tasks; creating a DSL by discovering
and reusing domain-specific subroutines; and training a neural network
that efficiently guides search for programs in the DSL. %% The learned DSL
%% distills commonalities across programs that solve tasks, helping
%% the agent solve related program induction problems. The neural
%% network ensures that searching for programs remains tractable
%% even as the DSL (and hence the search space for programs) expands.

The learned DSL effectively encodes a prior on programs likely to
solve tasks in the domain, while the neural net looks at the example
input-output pairs for a specific task and produces a ``posterior''
for programs likely to solve that specific task.  The neural network
thus functions as a \textbf{recognition model} supporting a form of
approximate Bayesian program induction, jointly trained with a
\textbf{generative model} for programs encoded in the DSL, in the
spirit of the Helmholtz machine~\cite{hinton1995wake}. The
recognition model ensures that searching for programs remains
tractable even as the DSL (and hence the search space for programs)
expands.

%% Our \system
%% algorithm incorporates these insights by iterating
Concretely, our algorithm iterates through three steps. 
The \textbf{Wake} step takes a given set of \textbf{tasks}, typically
several hundred, and searches for compact programs that solve these tasks
guided by the current DSL and neural network.
The \textbf{Sleep-G} step grows the DSL (or \textbf{G}enerative model) 
which allows the agent to more compactly
write programs in the domain; it modifies the structure of the DSL by
discovering regularities across programs found during waking, compressing
them to distill out common code fragments across successful programs.
The \textbf{Sleep-R} step improves the search procedure by training a neural network (the \textbf{R}ecognition model) to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}. %%  (CITE ,
%% and others too on amortized inference.)
Crucially, these steps bootstrap off each other~(Figure~\ref{feeding}).
\begin{figure}[h]\centering
  \begin{tikzpicture}
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs};
    \node[align = center](DSL) at (3,1) {DSL};
    \node[align = center](recognitionModel) at (9,1) {Neural\\recognition \\model};

    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
  \end{scope}
    \end{tikzpicture}
  \caption{Algorithm solves for programs, the DSL, and a recognition model. These steps bootstrap off each other in a Helmholtz-machine inspired wake/sleep inference strategy.}  \label{feeding}
\end{figure}

%% We apply \system to three domains:
%% list processing; text editing (in the style of FlashFill~\cite{gulwani2011automating}); and symbolic regression.
%%  For each of these we initially provide a generic
%%  set of programming primitives.
%% Our algorithm then constructs
%% its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).

Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}),
neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
which also learns a library of subroutines.
We find EC motivating, and go beyond it and other prior work through the following contributions: %% \begin{itemize}
%%   \setlength\itemsep{0}
%%   \item We show how to learn-to-learn programs in an expressive Lisp-like programming language;
%%   \item We give an algorithm for learning DSLs, built on a formalism known
%%     as Fragment Grammars~\cite{tim}; and 
%%   \item We give a hierarchical Bayesian framing of the problem that allows a
%%     joint inference of the DSL and recognition model.
%% \end{itemize}
\\\textbf{Contributions.} (1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
(2) We give an algorithm for learning DSLs,
built on a formalism known as Fragment Grammars~\cite{tim};
and (3) We give a hierarchical Bayesian framing of
the problem that
allows joint inference of the DSL and neural recognition model.

@@

@@

\begin{figure}\centering
  \begin{tabular}{lll}
    \toprule
    Name & Input & Output \\\midrule
    repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
    rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
    keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
    product & [7\, 1\, 6\, 2] & 84 \\
    \bottomrule
  \end{tabular}
  \captionof{table}{Some tasks in our list function domain.}\label{listexamples}
\end{figure}
We created 236 human-interpretable list manipulation tasks, each with 15
input/output examples (Tbl.~\ref{listexamples}).
%% Our data set is interesting in three major ways: many of the tasks
%% require complex solutions; the tasks were not generated from some
%% latent DSL, and the agent must learn to solve these complicated
%% problems from only 236 tasks.
Our data set assumes arithmetic operations as well as sequence operations,
so we additionally provide our system with the following arithmetic
primitives: \code{mod}, \code{*}, \code{>}, \code{is-square},
\code{is-prime}.

We evaluated \system on random 50/50 test/train split.  The system
composed 38 new subroutines, and rediscovered the higher-order
function \code{filter}, yielding a more expressive DSL more closely
matching the domain (left of Tbl.~\ref{initialExampleDSL}, right of
Fig.~\ref{fragmentExample}). See the supplement for a complete list of
DSL primitives discovered by \systemEnding.

@@

%% Because our enumerative search procedure cannot generate string %
%% constants, we instead enumerate programs with string-valued
%% parameters.  For example, to learn a program that prepends ``Dr.'', we
%% enumerate $\text{\code{(}}f_3\code{ string s)}$ -- where $f_3$ is the
%% learned appending primitive (Fig.~\ref{initialExampleDSL}) --- and then
%% define $\probability[x|p]$ by approximately marginalizing out the
%% string parameters via a simple dynamic program.
%% In Sec.~\ref{regressionSection}, we will use a similar trick to
%% synthesize programs containing real numbers, but using gradient
%% descent instead of dynamic programming.

@@

\section{Symbolic Regression: Programs from visual input}\label{regressionSection}
We apply \system
to symbolic regression problems.  Here, the
agent observes points along the curve of a function, and must write a
program that fits those points.  We initially equip our learner with
addition, multiplication, and division, and task it with solving
100 symbolic regression problems, each either a polynomial of degree
1--4 or a rational function.  The recognition model is a
convolutional network that observes an image of the target function's
graph (Fig.~\ref{functions}) --- visually, different kinds of
polynomials and rational functions produce different kinds of graphs,
and so the recognition model can learn to look at a graph and predict
what kind of function best explains it.  %% A key difficulty, however, is
%% that these problems are best solved with programs containing real
%% numbers.  Our solution to this difficulty is to enumerate
%%  programs with real-valued parameters, and then fit those
%% parameters by automatically differentiating through the programs the
%% system writes and use gradient descent to fit the parameters.
%% We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
%% and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.

\begin{figure}\vspace{-0.0cm} \newcommand{\functionSize}{1.5cm}\centering
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/48.png}
  \includegraphics[width = \functionSize]{figures/functions/102.png}
  \includegraphics[width = \functionSize]{figures/functions/116.png}\\
  \vspace{1pt}\includegraphics[width = \functionSize]{figures/functions/181.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/149.png}
    \includegraphics[width = \functionSize]{figures/functions/187.png}
  \caption{Recognition model input for symbolic regression. DSL learns subroutines for polynomials (top row) and rational functions (bottom) while the recognition  model jointly learns to look at a graph of the function (above) and predict which learned subroutines best explain the observation.}\label{functions}
\end{figure}
\system learns a DSL containing 13 new functions,
most of which are templates for polynomials of different orders or ratios of polynomials.
It also learns to find programs that minimize the number of continuous degrees of freedom.
For example, it learns to represent linear functions with the program
\code{(* real (+ x real))}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing --- both the implicit bias towards shorter programs and the likelihood model's BIC penalty.

\section{Learning from Scratch}
A long-standing dream within the program induction community
is ``learning from scratch'': starting with a \emph{minimal} Turing-complete programming language,
and then learning to solve a wide swath of
induction problems~\cite{solomonoff1964formal}.
All existing systems,
including ours,
fall far short of this dream,
and it is unclear (and we believe unlikely)
that this dream could ever be realized in its fullness.
How far can our system push in this direction?

``Learning from scratch'' is subjective, but we believe a reasonable
starting point is the set of primitives provided in McCarthy's 1959
Lisp implementation~\cite{mccarthy1960recursive}: these include
conditionals, \code{lambda}'s, recursion, arithmetic, and the basic
Lisp operators \code{cons}, \code{car}, \code{cdr}, and \code{nil}.
A sensible and modest goal is to start with
1959 Lisp primitives,
and then recover a DSL that
more closely resembles modern functional languages like Haskell and OCaml.
Recall (Section~\ref{sequences})
that we initially provided our system with modern functional programming routines like
\code{map} and \code{fold}.

We ran the following experiment: \system was provided with a subset of
the 1959 Lisp primitives, and tasked with solving 22 functional
programming exercises.  After running for 93 hours on 64 CPUs, our
algorithm solves these exercises, along the way assembling a DSL
containing 9 subroutines, recovering a modern repertoire of
functional programming idioms and subroutines, including \code{map},
\code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length},
and simple arithmetic operations like incrementing, decrementing, and
building lists of natural numbers between an interval.
We did not use the recognition model for this experiment:
a bottom-up pattern recognizer is of little use
for acquiring this abstract knowledge
from less than two dozen problems.

To be clear, we believe that program learners should \emph{not}
start from scratch,
but instead should start from
a rich, domain-agnostic
basis like those embodied in the standard libraries of modern programming languages.
What this experiment shows is that \system doesn't \emph{need} to start from a rich basis,
and can in principle recover many of the amenities of modern programming systems,
provided it is given enough computational power and a suitable
spectrum of programming exercises.

@@

We compare with ablations of our model on held out tasks.
The purpose of this ablation study is 
both to examine the role of each component of \systemEnding,
as well as to compare with
prior approaches in the literature:
a head-to-head
comparison of program synthesizers is complicated by the fact that
each system, including ours, makes idiosyncratic 
assumptions about the space of programs and the statement of tasks.

Nevertheless, much prior work can be modeled within our setup. 
We compare with the following ablations (Tbl~\ref{baselineComparisons};
Fig~\ref{learningCurves}):
\\\noindent \textbf{No NN:} lesions the recognition model.
\\\noindent \textbf{NPS}, which does not learn the DSL,
instead learning the recognition model
from samples drawn from the fixed DSL.
We call this NPS (Neural Program Synthesis)
because this is closest to how
RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
\\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL. This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
\\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first search step.
%For each domain,
%% We are interested both in how many tasks the
%% agent can solve and how quickly it can find those solutions.
%% Tbl.~\ref{baselineComparisons}
%% compares our model against these alternatives.
%% We consistently
%% improve on the baselines,
%% and find that lesioning the recognition model
%% and lesioning it also slows down the convergence of the algorithm,
%% taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
%% This supports a view of the recognition model as a way of amortizing the cost of search.
